Modeling and analyzing the Corona-virus warning app with the Isabelle Infrastructure framework

### 2020-07-11 submitted_version_paper.tex
new (submitted) version: fixed citation

%The CWA
%has taken a long time to develop being published only on
%16th June 2020. It
%was a quite costly project but this was mainly due to the management of
%Telekom and SAP being in the driving seat. But the app has been designed with great
The app was designed with great
attention on privacy: a distributed architecture \cite{cwa:arch} has been adopted after a long and
heated debate with supporters of a central architecture. The distributed architecture is
based on a very clever application design whereby users' phones broadcast
highly anonymized so called ``Ephemeral IDs'' at physical locations via the
Bluetooth Low Energy Beacon protocol.
The app saves those IDs of people in close proximity.
When at a later date an infected person reports his infection to a central server,
all clients are provided the means to reconstruct the IDs used by this particular
person's device over the last 14 days.
Therefore, any client can evaluate exposure risks locally and stored contact data has never to be shared.
%
%, the unique root ID is published
%and in the daily check all mobile phones connecting to the central server download
%the root IDs of infected people. Since the Ephemeral IDs can be mapped to the root ID
%all Ephemeral IDs that have been saved over the last 14 days allow users' phones to
%regularly check whether their user has been exposed to an infected person and issue
%a warning to the user and recommendation to contact health authorities.
%The warning issued by the Corona warning app entitles to
%having a Corona test done (which at the time of writing is not normally possible).

@@

The app now in use in Germany has been developed based on a protocol proposed by the DP-3T project \cite{dp3t:github}:
A sophisticated security concept
conceived by experts in the field that has strong claims with regard to mathematical support \cite{dp3t:wp} [p2].
%The Corona-virus app has been produced based on a sophisticated security
%concept conceived by experts in the field %to our knowledge, no formal verification
%%
%and has strong claims with regard to mathematical support:
%
%``Such a design builds on strong, mathematically provable support for privacy and data protection goals [...]'' \TODOref{DP-3T Whitepaper p. 2} \smallskip\\
%  \NOTE{DP-3T on their design and how wisely chosen it is ;) We have to use this! Something along the l%ines of:}
%
However, there has, as of yet, to our knowledge no formal verification been involved.
Even if a ``post-production'' formal specification seems pointless, 
it allows to reveal weak points of the architecture, show that the measures that have
been conceived are suitable to cover those weak points,
or where and to what extent trade offs have to be made due to inherent vulnerabilities.
The protocol implemented by the framework CWA resembles only the most basic
of the three proposed by the DP-3T project.
Hence, a formal verified model that stresses the impact or limits of certain security measures
might give more weight to appeals like \cite{dp3t:impl} 
%to Apple and Google
to adopt the more sophisticated protocols.
%
%Thereby, we believe
%that our current work is useful to increase the trust in the Corona-virus app necessary for
%its wide adoption which in turn is crucial for it to be efficient.

@@

In this paper, we first provide some background in Section \ref{sec:background}:
we give a detailed overview of the development history %and security and privacy
and relevant parts of the CWA (Section \ref{sec:history})
We then introduce the Isabelle Infrastructure framework (Section \ref{sec:isainf}).
Next, we present our model (Section \ref{sec:model}) and analysis of privacy and
attacks (Section \ref{sec:ana}). The found attack on the first abstract specification
motivates refinement. The formal definition of refinement for the Isabelle Infrastructure
framework is introduced and illustrated on the CWA (Section \ref{sec:ref})
before drawing some conclusions (Section \ref{sec:concl}).

@@

\section{Background and related work}
\label{sec:background}
\subsection{DP-3T and PEPP-PT}
\label{sec:history}
We are mainly concerned with the architecture and protocols proposed by the
DP-3T (\textit{Decentralized Privacy-Preserving Proximity Tracing}) project \cite{dp3t:github}.
The main reason to focus on this particular family of protocols is the
\textit{Exposure Notification Framework} (ENF) \cite{enf:proj}, jointly published by Apple and Google, that adopts
%the main
core principles of the DP-3T proposal. This API is not only used in the German
CWA but has the potential of being widely adopted in future
app developments that might emerge due to the reach of players like Apple and Google.

There are, however, competing architectures noteworthy, namely protocols developed under the
roof of the \textit{Pan-European Privacy-Preserving Proximity Tracing} project (PEPP-PT) \cite{pepppt:github}, e.g.
PEPP-PT-ROBERT \cite{pepppt:ROBERT}.

Neither DP-3T nor PEPP-PT are synonymous for just one single protocol. Each project endorses
different protocols with unique properties in terms of privacy and data protection.
Yet, on a higher level of abstraction, it seems feasible to distinguish two basic architectures:
protocols as endorsed by PEPP-PT might be characterized as centralized architectures whereas
DP-3T-inspired protocols follow a (more) decentralized approach.\footnote{%
  DP-3T involves a central backend server. It is decentralized with regard to
  the collection and evaluation of contact information:
  In centralized architectures the server provides a risk scoring services, whereas decentralized
  approaches rely on local risk assessment and, thus, do not need to share contact information with
  the backend.}
\\

All DP-3T protocols roughly work the same:
%\subsubsection{Basic DP-3T protocol: How does it work?}
%  How does it relate to ENF/CWA? (broad strokes)}
%
Upon installation, the app generates secret daily seeds to derive \textit{Ephemeral Ids}
(EphIDs) from them. EphIDs are generated locally with cryptographic methods and cannot be connected
to one another but only reconstructed from the secret seed they were derived from.

@@

In the most simple (and insecure) protocol proposed by DP-3T %\cite{dp3t:wp}[p14ff]
this basically translates into publishing the daily seeds used to derive EphIDs from.
The protocol implemented by ENF and,
hence, CWA adopts this low-cost design \cite{cwa:arch}. DP-3T proposes two other, more sophisticated protocols that improve
privacy and data protection properties to different degrees but are more costly to implement.
%Aside from additional security measures,
%\TODOfn[styC]{In short: I want to name it but don't talk about it (since we're not really concerned with% this level of detail). But this sounds just sh...}
%like signing content the server provides as a general rule, the protocol implemented by ENF and,
%hence, CWA follows this low-cost design.\TODOref{CWA solution architecture ($\rightarrow$ github:CWA); probably ENF specification} DP-3T proposes two other, more sophisticated protocols that improve
%privacy and data protection properties to different degrees but are more costly.
%%TODOfn[styC]{``complicated'' sounds lame but ``costly'' is, perhaps, to simple/literal} to set up.

Figure \ref{fig:dp3tprot} illustrates the basic system architecture
along with some of the mitigation measures either implemented in CWA or proposed by DP-3T.

\begin{figure}[htb]
\vspace{-.5cm}
  \begin{center}
    \includegraphics[width=5in]{DP-3T_basic_protocol}
  \end{center}
\vspace{-.5cm}
  \caption{Decentralized privacy-preserving proximity tracing protocol of CWA}
  \label{fig:dp3tprot}
\vspace{-.5cm}
\end{figure}

>>>>

%The CWA
%has taken a long time to develop being published only on
%16th June 2020. It
%was a quite costly project but this was mainly due to the management of
%Telekom and SAP being in the driving seat. But the app has been designed with great
The app was designed with great
attention on privacy: a distributed architecture \cite{cwa:arch} has been adopted %after a long and
%heated debate with supporters of a central architecture. The distributed architecture is
that is
based on a very clever application design whereby clients broadcast
highly anonymized so called ``Ephemeral Ids'' via Bluetooth and
store those ids of people in close proximity.
%highly anonymized so called ``Ephemeral IDs'' at physical locations via the
%Bluetooth Low Energy Beacon protocol.
%The app saves those ids of people in close proximity.
Infected persons report their infection by uploading their ids to a central server, providing all clients the means to assess exposure risk locally,
hence, stored contact data has never to be shared.
%Infected persons report their infection to a central server,
%which, in turn, provides all clients the means to reconstruct the ids
%used by reported cases over the last 14 days.
%
%When at a later date an infected person reports his infection to a central server,
%all clients are provided the means to reconstruct the ids used by this particular
%person's device over the last 14 days.
%Therefore, exposure risks can be evaluated locally and stored contact data has never to be shared.
%
%, the unique root ID is published
%and in the daily check all mobile phones connecting to the central server download
%the root IDs of infected people. Since the Ephemeral IDs can be mapped to the root ID
%all Ephemeral IDs that have been saved over the last 14 days allow users' phones to
%regularly check whether their user has been exposed to an infected person and issue
%a warning to the user and recommendation to contact health authorities.
%The warning issued by the Corona warning app entitles to
%having a Corona test done (which at the time of writing is not normally possible).

@@

The app now in use in Germany has been developed based on a protocol proposed by the DP-3T project \cite{dp3t:github}:
A sophisticated security concept
conceived by experts in the field that has strong claims with regard to mathematical support \cite{dp3t:wp} [p2].
%
However, there has, as of yet, to our knowledge no formal verification been involved.
Even if a ``post-production'' formal specification seems pointless, 
it allows to reveal weak points of the architecture, show that the measures that have
been conceived are suitable to cover those weak points,
or to what extent trade offs have to be made due to inherent vulnerabilities.
The protocol implemented by the framework CWA resembles only the most basic
of the three protocols proposed by the DP-3T project.
Hence, a formal verified model that stresses the impact or limits of certain security measures
might give more weight to appeals like \cite{dp3t:impl}
to adopt the more sophisticated protocols.

@@

In this paper, we first provide some background in Section \ref{sec:background}:
we give a brief overview of related works
and the protocol of the German app (Section \ref{sec:history}).
We then introduce the Isabelle Infrastructure framework (Section \ref{sec:isainf}).
Next, we present our model (Section \ref{sec:model}) and analysis of privacy and
attacks (Section \ref{sec:ana}). The found attack on the first abstract specification
motivates refinement. The formal definition of refinement for the Isabelle Infrastructure
framework is introduced and illustrated on the CWA (Section \ref{sec:ref})
before drawing some conclusions (Section \ref{sec:concl}).

@@

\section{Background and related work}
\label{sec:background}
\subsection{DP-3T and PEPP-PT}
\label{sec:history}
We are mainly concerned with the architecture and protocols proposed by the
DP-3T (\textit{Decentralized Privacy-Preserving Proximity Tracing}) project \cite{dp3t:github}.
The main reason to focus on this particular family of protocols is the
\textit{Exposure Notification Framework} (ENF), jointly published by Apple and Google \cite{enf:proj}, that adopts
%the main
core principles of the DP-3T proposal. This API is not only used in the German
CWA but has the potential of being widely adopted in future
app developments that might emerge due to the reach of players like Apple and Google.

There are, however, competing architectures noteworthy, namely protocols developed under the
roof of the \textit{Pan-European Privacy-Preserving Proximity Tracing} project (PEPP-PT) \cite{pepppt:github}, e.g.
PEPP-PT-ROBERT \cite{pepppt:ROBERT},
that might be characterized as centralized architectures.

Neither DP-3T nor PEPP-PT are synonymous for just one single protocol. Each project endorses
different protocols with unique properties in terms of privacy and data protection.
%Yet, on a higher level of abstraction, it seems feasible to distinguish two basic architectures:
%protocols as endorsed by PEPP-PT might be characterized as centralized architectures whereas
%DP-3T-inspired protocols follow a (more) decentralized approach.\footnote{%
%  DP-3T involves a central backend server. It is decentralized with regard to
%  the collection and evaluation of contact information:
%  In centralized architectures the server provides a risk scoring services, whereas decentralized
%  approaches rely on local risk assessment and, thus, do not need to share contact information with
%  the backend.}

There is a variety of noteworthy privacy and security issues. %that deserve formal consideration (and verification).
The debate among advocates of centralized architectures and those in favor of a decentralized approach in particular yields a lot of interesting material detailing different attacks and possible mitigation strategies: \cite{pepppt:dp3tana}, \cite{pepppt:dp3tresp}, \cite{dp3t:psre}.

In terms of attack scenarios, we focus on, what might be classified as deanonymisation attacks: Tracking a device (see \cite{dp3t:psre}[p9], \cite{pepppt:dp3tana}[p8]) and identifying infected individuals (\cite{dp3t:psre}[p5], \cite{pepppt:dp3tana}[p9]).

\subsubsection{Basic DP-3T architecture}
Upon installation, the app generates secret daily seeds to derive \textit{Ephemeral Ids}
(EphIDs) from them. EphIDs are generated locally with cryptographic methods and cannot be connected
to one another but only reconstructed from the secret seed they were derived from.

@@

In the most simple (and insecure) protocol proposed by DP-3T %\cite{dp3t:wp}[p14ff]
this basically translates into publishing the daily seeds used to derive EphIDs from.
The protocol implemented by ENF and,
hence, CWA adopts this low-cost design \cite{cwa:arch}. DP-3T proposes two other, more sophisticated protocols that improve
privacy and data protection properties to different degrees but are more costly to implement.
%
Figure \ref{fig:dp3tprot} illustrates the basic system architecture
along with some of the mitigation measures either implemented in CWA or proposed by DP-3T.
\begin{figure}[htb]
\vspace{-.5cm}
  \begin{center}
    \includegraphics[width=5in]{DP-3T_basic_protocol}
  \end{center}
\vspace{-.5cm}
  \caption{Decentralized privacy-preserving proximity tracing protocol of CWA}
  \label{fig:dp3tprot}
\vspace{-.5cm}
\end{figure}
